---
title: "PA1_template"
author: "Michael Kolb"
date: "December 10, 2017"
output:
  html_document: default
  pdf_document: default
---

Load and process the data

```{r}
data <- read.csv("activity.csv")
data$date <- as.Date(data$date)
cdata <- data[!is.na(data$steps),]
```

Calculate the total number of steps taken per day

```{r}
sumSteps <- tapply(cdata$steps, cdata$date, sum)
```

Make a histogram of the total number of steps taken each day

```{r}
hist(sumSteps, xlab="Steps", main="Total Steps per Day")
```

Report mean and median of the total number of steps taken per day

```{r}
mean(sumSteps)
median(sumSteps)
```

Make a time series plot of the 5-min interval and the average number of steps

```{r}
avgSteps <- aggregate(cdata$steps, by=list(cdata$interval), FUN=mean)
names(avgSteps) <- c("Interval", "Mean")
plot(avgSteps$Interval, avgSteps$Mean, type="l", xlab="Interval", 
	ylab="Average Number of Steps", main="Average Steps per Interval")
```

Identify minute interval that contains maximum number of steps

```{r}
avgSteps[which.max(avgSteps$Mean), ]$Interval
```

Calculate and report the total number of missing values in the dataset

```{r}
sum(is.na(data$steps))
```

Create a new dataset equal to original dataset with missing data filled in

```{r}
idata <- avgSteps$Mean[match(cdata$interval, avgSteps$Interval)]
imputedSteps <- transform(data, steps= ifelse(is.na(data$steps), yes= idata,
	no= data$steps))
totalSteps <- aggregate(steps ~ date, imputedSteps, sum)
names(totalSteps) <- c("Date","Steps")
```

Make a histogram of the total number of steps taken each day

```{r}
hist(totalSteps$Steps, xlab="Steps", main="Total Steps per day")
```

Report mean and median total number of steps taken per day

```{r}
mean(totalSteps$Steps)
median(totalSteps$Steps)
```

Create factor variable with two levels indicating weekday or weekend day

```{r}
totSteps <- aggregate(steps ~ interval + date, imputedSteps, sum)
totSteps$dayType <- as.factor(sapply(totSteps$date, function(x) {
	if (weekdays(x)=="Saturday" | weekdays(x)=="Sunday")
	y <- "Weekend" else
	y <- "Weekday"
	y
}))
```

Make panel plot of 5-min interval & avg steps across weekday / weekend days

```{r}
par(mfrow= c(2,1))
with(totSteps[totSteps$dayType=="Weekend",], plot(aggregate(steps ~ interval,
	FUN=mean), type="l", main="Weekend Activity Pattern"))
with(totSteps[totSteps$dayType=="Weekday",], plot(aggregate(steps ~ interval,
	FUN=mean), type="l", main="Weekday Activity Pattern"))
```